Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-07-09 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

Mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin FEM is used to obtain high-order fluxes on CV boundaries. Smooth transition between first- and high-order fluxes are enforced through an extrema detection scheme based on normalised variable diagram \citep[NVD,][]{jasak_1999,darwish_2003} applied across faces of neighbouring CVs. Resulting normalised upwind face values are used to calculate flux-limited solutions \citep[based on total variation diminishing criteria, TVD,][]{piperno_1998} to yield bounded fields \citep[\eg positive densities, saturations bounded between 0 and 1 etc. See][for more details on the flux-limiting scheme within the model]{gomes_book_2012}. Simulations performed for this work were conducted using three types of elements: \PN[1]{1}, \PN[1]{2} and \PNDG[1]{1} (Section~\ref{section:results_3D}).

@@

\medskip
Here, 2D simulations were conducted in a 5$\times$5 cm domain (Fig.~\ref{fig:homoheleshaw_VN3}a) fully saturated with a fluid. Wetting phase fluid is driven from the bottom left-hand corner of the domain with velocity of 1 cm.s$^{-1}$ (magnitude). No-flux boundary conditions were imposed to all borders of the domain except at the top right-hand side corner (named as sink), and a pressure gradient between source and sink regions was initially imposed to the system.

@@

The dynamic mesh adaptive algorithm embedded in the Fluidity/IC-FERST model utilises a metric tensor field dependent on solution interpolation error-estimates which locally control the topology of elements in the metric \citep{pain_2001}. Mesh optimisation generates unstructured finer mesh in regions where flow properties change faster, and coarser mesh in regions where properties change more slowly \citep{piggott_2006,hiester_2014}. %The mesh adapts in three stages: metric formation, mesh optimisation and fields' interpolation from the pre- to post-adapted mesh \citep[see][]{hiester_2014}. 
In the simulations shown in this work, mesh will adapt as a response to oscillations in the phase saturation field with prescribed interpolation error estimate \citep{mostaghimi_2016}.

>>>>

Mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin FEM is used to obtain high-order fluxes on CV boundaries. Smooth transition between first- and high-order fluxes are enforced through an extrema detection scheme based on normalised variable diagram \citep[NVD,][]{jasak_1999,darwish_2003} applied across faces of neighbouring CVs. Resulting normalised upwind face values are used to calculate flux-limited solutions \citep[based on total variation diminishing criteria, TVD,][]{piperno_1998} to yield bounded fields \citep[\eg positive densities, saturations bounded between 0 and 1 etc. See][for more details on the flux-limiting scheme within the model]{gomes_book_2012}. Simulations performed for this work were conducted using three types of elements: \PN[1]{1}, \PN[1]{2} (Sections~\ref{section:results_homo_hete} and~\ref{section:results_hete_fix_adapt}) and \PNDG[1]{1} (Section~\ref{section:results_3D}).

@@

\medskip
Here, 2D simulations were conducted in a 5$\times$5 cm domain (Fig.~\ref{fig:homoheleshaw_VN3}a) fully saturated with a fluid. Wetting phase fluid is driven from the bottom left-hand corner of the domain with velocity of 1 cm.s$^{-1}$ (magnitude). No-flux boundary conditions were imposed to all borders of the domain except at the top right-hand side corner (named as sink), and pressure gradient between source and sink regions was initially imposed to the system. Solution mesh-independence was achieved through comparison of non-wetting phase saturation profiles along a line between source and sink regions (diagonal across the domain) with several mesh resolutions. Mesh convergence criterion was a maximum residual of 10$^{-2}$ and it was achieved with approximately 3.5k elements (Fig.~\ref{fig:MeshDependence}). Thus, all simulations shown in this Section used mesh with resolution larger than 3546 \PN[1]{2} triangular elements.

@@

The dynamic mesh adaptive algorithm embedded in the Fluidity/IC-FERST model utilises a metric tensor field dependent on solution interpolation error-estimates which locally control the topology of elements in the metric \citep{pain_2001}. Mesh optimisation generates unstructured finer mesh in regions where flow properties change faster, and coarser mesh in regions where properties change more slowly \citep{piggott_2006,hiester_2014,mostaghimi_2016}. %The mesh adapts in three stages: metric formation, mesh optimisation and fields' interpolation from the pre- to post-adapted mesh \citep[see][]{hiester_2014}. 
In the simulations shown in this section, mesh grid adapts as a response to oscillations of phase saturations with: (a) minimum value for length of elements of 0.05 unit-length, (b) maximum number of nodes of 5$\times$10$^{4}$, and (c) prescribed interpolation error estimate of 10$^{-2}$.

### 2018-07-09 Papers/ViscousInstabilities/article_figure1.tex
updates

%%%%
%%%%  FIGURE 
%%%%
\begin{figure}[h]
\centering
%\vbox{\includegraphics[width=0.75\textwidth]{./Pics/phase_vol_frac_uni_perm_1.pdf}}
\vbox{\includegraphics[width=0.75\textwidth]{./Pics/phase_vol_frac_uni_perm_2.pdf}}
\caption{Schematics of formation of flow instabilities during injection of a pure low viscosity fluid (red) into a domain saturated with a second fluid (dark blue). The viscocity ratio of the two fluids is VR=$5$. In this case, the initially piston shape front collapses leading to the formation of several fingers.}
\label{fig:simple_case}
\end{figure}
\clearpage

@@

>>>>

%%%%
%%%%  FIGURE 
%%%%
%\begin{figure}[h]
%\centering
%\vbox{\includegraphics[width=0.75\textwidth]{./Pics/phase_vol_frac_uni_perm_1.pdf}}
%\vbox{\includegraphics[width=0.75\textwidth]{./Pics/phase_vol_frac_uni_perm_2.pdf}}
%\caption{Schematics of formation of flow instabilities during injection of a pure low viscosity fluid (red) into a domain saturated with a second fluid (dark blue). The viscocity ratio of the two fluids is VR=$5$. In this case, the initially piston shape front collapses leading to the formation of several fingers.}
%\label{fig:simple_case}
%\end{figure}
%\clearpage

@@

%%%%
%%%%  FIGURE 
%%%%
\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.75\textwidth]{./Pics1/Saffman_homogeneous/MeshIndependence.pdf}
  \caption{Mesh convergence analysis for Hele-Shaw cells flow simulations: non-wetting phase saturation profiles through a center-line across the domain (diagonal) with grid resolution ranging from 1881 to 3546 \PN[1]{2} triangular elements.}
\end{center}
\label{fig:MeshDependence}
\end{figure}
\clearpage
